
---
title: "Code examples in R"
subtitle: "Translated code examples from In-Memory Analytics with Apache Arrow by Matthew Topol"
author: François Michonneau
format:
  html:
    code-fold: false
    footnotes-hover: true
    df-print: kable
    toc: true
---

## Chapter 1

### Setting up your shooting range

This will work for most people. If you are using Linux, you have more
options avavailable to you that are highlighted in a
[vignette](https://arrow.apache.org/docs/r/articles/install.html) that
comes with the `{arrow}` package.

```{r}
#| eval: false

install.package("arrow")
```

```{python}
import pyarrow as pa
import numpy as np
NROWS = 8192
NCOLS = 16
data = [pa.array(np.random.randn(NROWS)) for i in range(NCOLS)]
cols = ['c' + str(i) for i in range(NCOLS)]
rb = pa.RecordBatch.from_arrays(data, cols)
print(rb.schema)
print(rb.num_rows)
```

```{r}
library(arrow)
NROWS <- 8192
NCOLS <- 16
data <- replicate(
  NCOLS,
  rnorm(NROWS),
  simplify = FALSE
) |>
  as.data.frame(
    col.names = paste0("c", seq_len(NCOLS))
  )
rb <- record_batch(data)
rb
nrow(rb)
```

### Building a struct array

```{python}
archer_list = [{
    'archer': 'Legolas',
    'location': 'Murkwood',
    'year': 1954,
},{
    'archer': 'Oliver',
    'location': 'Star City',
    'year': 1941,
},{
    'archer': 'Merida',
    'location': 'Scotland',
    'year': 2012,
},{
    'archer': 'Lara',
    'location': 'London',
    'year': 1996,
},{
    'archer': 'Artemis',
    'location': 'Greece',
    'year': -600,
}]

archer_type = pa.struct([('archer', pa.utf8()),
                         ('location', pa.utf8()),
                         ('year', pa.int16())])
archers = pa.array(archer_list, type=archer_type)
print(archers.type)
print(archers)
```

```{r}
archer_list <- tibble::tribble(
  ~ archer,  ~ location,  ~ year,
  "Legolas", "Murkwood",  1954,
  "Oliver",  "Star City", 1941,
  "Merida",  "Scotland",  2012,
  "Lara",    "London",    1996,
  "Artemis", "Greece",    -600
)

archer_type <- struct(archer = utf8(), location = utf8(), year = int16())
archers <- Arrow$create(
  archer_list,
  type = archer_type
)
archers$type
archers
```

There is not a direct equivalent between a Python dictionary in
R. Given that Arrow works with rectangular data, we use the same data
formatted as a tibble (constructed with the `tribble()` function for
easier visual representation) as input.

In R, Arrow struct arrays are created automatically based on the
dimensions of the object provided to the constructor. Atomic vectors
are converted into Arrow arrays, while data frames (and by extension
tibbles) are converted into Arrow struct arrays.

Instead of using the R6 constructor `Arrow$create()`, we could also
have used the `as_arrow_array()` function. Note
[issue](https://issues.apache.org/jira/browse/ARROW-17620).

